---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

This step is completely straightforward.  I unzip the existing file in the archive (no need to check in the csv),
and read the frame, which already has column headers. 
```{r}
unzip("activity.zip")
activityData<-read.csv("activity.csv")
dim(activityData)
head(activityData,n=1)
library(plyr)
library(ggplot2)
```

## What is mean total number of steps taken per day?

```{r}
totals<-ddply(activityData,.(date),summarise,steps=sum(steps,na.rm=F))
meanSteps<-mean(totals$steps,na.rm=T)
meanSteps
```

## What is the average daily activity pattern?

We collect the mean number steps by interval over all dates given (removing the NA row/column combos),
and we get:

```{r}
intervalTotals<-ddply(activityData,c("interval"),.fun=function(x) mean(x$steps,na.rm=T))
dim(intervalTotals)
head(intervalTotals)
q<-ggplot(data=intervalTotals,aes(x=interval,y=V1))+geom_line()
q
```

Almost no steps are taken in the intervals from 0 to 500, and the number of steps peaks at around interval 800.

## Imputing missing values  using mean for day and interval


```{r}
imputeRow<-function(date,interval){
  mean(activityData$steps[activityData$interval==interval&activityData$date==date],na.rm = T)
}

activityDataFilled<-transform(activityData,steps=ifelse(is.na(steps),imputeRow(date,interval),steps))

head(activityDataFilled)

```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
activityDataFilled<-transform(activityDataFilled,weekday=weekdays(as.Date(date)))
weekdaySummary<-ddply(activityDataFilled,.(weekday),total=mean(steps))
q<-ggplot(data=activityDataFilled,aes(x=weekday,y=steps))+geom_point()
q
```

The averages are the same, but activity seems more bi-modal on weekends, with many people taking fewer steps and a few taking far more steps, and few in between. 
